---
title: "Ch. 10 Relational Data pg 171-193"
author: "Ann-Marie Zahn"
date: "October 24, 2017"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(nycflights13)
library(maps)
```

###Exersizes 1: pg 174-175, #'s 1-4

######1.

You want to draw the route each plane flies from its origin to its destination.

Necesary Variables: origin, destination, latitude, longitude 

Necesary Tables to combine: airports, planes

######2.

Relationship between weather and airports: weather and airports share origin. They should be connected.

######3.

If weather had info for all airports you would also need a relation with destination.

######4.

Special table would include: month, day

Primary key would have to be both month and day together

Connections: special would connect to flights (month, day) and weather (month, day)

###Excersize 2: pg. 177, #'s 1-3

######1. 

```{r ex2.1}

#add surrogate key to flights (create primary key)
flights <- flights %>%
  mutate(Seq = row_number())

```

######2.

```{r ex2.2, eval=FALSE}

Lahman::Batting

babynames::babynames

nasaweather::atmos

fueleconomy::vehicles

ggplot2::diamonds

Lahman::TeamsFranchises

```

Lahman::Batting

* primary key: playerID
* foreign key: teamID, yearID, teamID, lgID

babynames::babynames

* primary key: name and year together
* foreign key: sex

nasaweather::atmos

* primary key: lat and long together
* foreign key: ozone, temp

fueleconomy::vehicles

* primary key: id
* foreign key: make, model

ggplot2::diamonds

* primary key: none
* foreign key: carat, cut, color, clarity, depth

######3. 

drawing!!!!!!!

###Excersize 3: pg. 186-187, #'s 1-5

######1. 

```{r ex3.1}

flights %>%
  group_by(dest) %>%
  summarise(avg = mean(arr_delay, na.rm = TRUE)) %>%
  left_join(airports, c(  "dest" = "faa")) %>%
  ggplot(aes(lon, lat, color = avg)) +
  borders("state") +
  geom_point() +
  coord_quickmap()
    
```

######2. 

```{r ex3.2}

flights2 <- flights %>%
  left_join(airports, c(  "dest" = "faa"), suffix = c(".dest", ".origin")) %>%
  left_join(airports, c("origin" = "faa"), suffix = c(".dest", ".origin")) %>%
  select(-name.dest, -(alt.dest:name.origin), -(alt.origin:tzone.origin))

```

######3. 

```{r ex3.3}

flights %>%
  left_join(planes, by = "tailnum", suffix = c(".flight", ".plane")) %>%
  select(-(type:engine)) %>%
  ggplot(aes(x = year.plane, y = arr_delay)) + geom_jitter()

```

There does not seem to be a relationship between the age of a plane and its arrival delay. This may be because there are just not as many older planes still flying.

######4. 

```{r ex3.4}

flights3 <- flights %>%
  left_join(weather, by = c("time_hour", "origin"), suffix = c(".flight", ".weather"))

ggplot(data = flights3, aes(x = dep_delay)) + geom_jitter(aes(y = temp)) 

ggplot(data = flights3, aes(x = dep_delay)) + geom_jitter(aes(y=dewp))

ggplot(data = flights3, aes(x = dep_delay)) + geom_jitter(aes(y=humid))

ggplot(data = flights3, aes(x = dep_delay)) + geom_jitter(aes(y=precip))

ggplot(data = flights3, aes(x = dep_delay)) + geom_jitter(aes(y=visib))

ggplot(data = flights3, aes(x = dep_delay)) + geom_jitter(aes(y=wind_speed))

ggplot(data = flights3, aes(x = dep_delay)) + geom_jitter(aes(y=wind_gust))

```

It looks like temp, visibility, and percipitation all factor into the likelihood of a delay.

######5. 

```{r ex3.5}

flights %>%
  filter(year == 2013, month == 6, day == 13) %>%
  group_by(dest) %>%
  summarise(avg = mean(arr_delay, na.rm = TRUE)) %>%
  left_join(airports, c(  "dest" = "faa")) %>%
  ggplot(aes(lon, lat, color = avg)) +
  borders("state") +
  geom_point() +
  coord_quickmap()

```

The eastern coast expereinced a lot of delays due to two derechos, which contained high winds. 

###Excersize 4: pg. 191 #'s 1-6

######1. 

```{r ex4.1}

anti_join(flights, planes, by = "tailnum") %>%
  group_by(carrier) %>%
  count() 

```
It seems that mainly AA and MQ have tailnums that are not in the planes table.

######2. 

```{r ex4.2}

planes_100_plus_flights <- flights %>%
  group_by(tailnum) %>%
  count() %>%
  filter(n > 100)

semi_join(flights, planes_100_plus_flights, by = "tailnum")

```

######3. 

```{r ex4.3}

fueleconomy::vehicles %>%
  semi_join(fueleconomy::common)

```

######4. 

```{r ex4.4}



```

######5. 

```{r ex4.5}

anti_join(flights, airports, by = c("dest" = "faa"))

anti_join(airports, flights, by = c("faa" = "dest"))

```

`anti_join(flights, airports, by = c("dest" = "faa"))` gives us all the flight records for the destinations that are not in the airports table.

`anti_join(airports, flights, by = c("faa" = "dest"))` gives us all the airport records for airports that were not destinations in the flights table.

######6. 

```{r ex4.6}

flights %>%
  group_by(tailnum, carrier) %>%
  count() %>%
  group_by(tailnum) %>%
  count() %>%
  filter(nn > 1)

```

There is not an implicit relationship between a plane and airline. There are 17 planes that are flown by different airlines (not including NA). 